Supplemental Information:\\ Harnessing Design Principles from Glass Sponges for Structurally Robust Lattice Structures

### 2018-09-11 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
progress

\section{Designs Considered}
In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only. In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e. volume or mass) is conserved. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

\subsection{Cylindrical Cross-section}
For this cross-section type we stay consistent with the geometry found within the biological species. Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.

\subsubsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
respectively. Note that in this study we use 
\begin{equation}
\frac{D_{A,nd}}{L}=0.1,
\end{equation}
since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

Moreover, as for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell. As such, the diagonals intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes and their  volume and projected area in the unit cell are
\begin{equation}\label{V2}
V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
and
\begin{equation}\label{A2}
A_{A,d}=8\sqrt{2} L D_{A,d},
\end{equation}
respectively. Since the projected area ratio of the non-diagonal to diagonal elements in the sponge has been measured to be 
\begin{equation}
\frac{A_{A,nd}}{A_{A,d}}=1.4,
\end{equation}
by substituting \cref{A1} and \cref{A2} into the equation above we find that for Design A
\begin{equation} \label{DA}
D_{A,nd}=1.4\sqrt{2}D_{A,d}\approx 2 D_{A,d}.
\end{equation}
Substitution of \cref{DA} into \cref{V1} and \cref{V2} yields 
\begin{equation}
\frac{V_{A,nd}}{V_{A,d}}=\frac{2 L \pi D^2_{A,nd}}{2\sqrt{2}L \pi D^2_{A,d}}=2\sqrt{2}
\end{equation}
and
\begin{equation}\label{VT}
V_{A,T}=V_{A,nd}+V_{A,d}=2\pi L (D_{A,nd}^2+\sqrt{2} D_{A,d}^2)=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
where $V_{A,T}$ indicates the total volume of the unit cell for Design A. 
% such that \cref{VT} can be further simplified to
% \begin{equation}
% {V}_{A,T}=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right)\approx1.3288[\text{mm}^3].
% \end{equation}

@@

\subsubsection{Design B}
Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{DesignB}). However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. It follows that the non-diagonal and diagonal volumes are given by
\begin{equation}
V_{B,nd}=V_{A,nd}=2\pi L D_{B,nd}^2
\end{equation}
and
\begin{equation}
V_{B,d}=2\sqrt{8} L \left(\pi \frac{{D}_{B,d}^2}{4}\right),
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2}, as well as the above volumes, we  obtain 
\begin{equation}
{{D}_{B,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{B,d}}{{D}_{B,nd}}=\frac{1}{\sqrt{2}}.
\end{equation}

@@

\subsubsection{Design C}
Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{DesignC}). For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
\begin{equation}
V_{C,nd}=V_{A,nd}=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}
V_{C,d}=V_{A,d}=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2} we  obtain 
\begin{equation}
{{D}_{C,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{C,d}}{{D}_{C,nd}}=\frac{1}{2}.
\end{equation}

@@

\subsubsection{Design D} 
Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}. As such, for this design we allocate the total material volume to the non-diagonal elements. Since
\begin{equation}
V_{D,T}=V_{D,nd}=V_{A,nd}=2\pi L D_{D,nd}^2,
\end{equation}
using the constraint provided by \cref{con1} we obtain
\begin{equation}
{D_{D,nd}}={D_{A,nd}}\sqrt{1+\frac{\sqrt{2}}{4}}. 
\end{equation}

@@

\subsection{Cylindrical Cross-section Results}
\mf{add results and figures here}
Include graphs showing results for cylindrical cross-section, as can be found in the main paper.

\subsection{Polygonal Cross-section}
In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section. For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior. The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.

\subsubsection{Design A}
From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ . Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
\begin{equation}
	V_{A,nd}=8LD_{A,nd}t
\end{equation}
\begin{equation}
	V_{A,d}=8\sqrt{2}LD_{A,d}t
\end{equation}
Let us enter this into our biological observation, which yields:
\begin{equation}
	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
\end{equation}
Therefore, we can simplify this relationship as
\begin{equation}
	D_{A,nd}=2D_{A,d}
\end{equation}

\subsubsection{Design B}
For this design the volume of non-diagonal elements remain the same, namely,
\begin{equation}
		V_{B,nd}=8LD_{B,nd}t.
\end{equation}
However, the volume for the diagonal elements will is different as a result the change in total length, namely,
\begin{equation}
		V_{B,d}=4\sqrt{2}LD_{A,d}t.
\end{equation}
In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
\begin{equation}
		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
\end{equation}
which simplifies to
\begin{equation}
	D_{B,nd}=D_{B,d}
\end{equation}
and
\begin{equation}
	D_{B,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design C}
Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
\begin{equation}
D_{A,nd}=2D_{A,d}
\end{equation}
and
\begin{equation}
D_{C,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design D}
For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements. Thus we can formulate the total volume of Design A as
\begin{equation}
	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
\end{equation}
Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
\begin{equation}
V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}
\end{equation}

@@

\section{Domain Size Analysis}
In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value. This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.

@@

\section{Optimization Analysis Description}
For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}. The algorithm input variables are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction. Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements. For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.

>>>>

\section{Designs Considered}
In this study, we focus on four different lattice configurations constrained to deform in an in-plane setting only. In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e. volume or mass) is conserved. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry) and two different shapes for the cross section of all lattice members: circular and rectangular.  For circular cross-sections  we stay consistent with the geometry found within the biological species. However, since slender members with circular cross section easily buckle out-of-plane when the structure is subjected to compression, to facilitate  fabrication and testing and to avoid out-of-plane buckling, we also consider a rectangular cross-section. \KB{how do you determine the aspect ratio of the rectangular cross section? I don't get the text that follows from here}

For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior. The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.

In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section. For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior. The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.

In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

%For this cross-section type we stay consistent with the geometry found within the biological species. Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.

\subsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). 
\subsubsection{Circular cross section} 
If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
respectively. Note that in this study we use 
\begin{equation}
\frac{D_{A,nd}}{L}=0.1,
\end{equation}
since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

Moreover, as for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell. As such, the diagonals intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes and their  volume and projected area in the unit cell are
\begin{equation}\label{V2}
V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
and
\begin{equation}\label{A2}
A_{A,d}=8\sqrt{2} L D_{A,d},
\end{equation}
respectively. Since the projected area ratio of the non-diagonal to diagonal elements in the sponge has been measured to be 
\begin{equation}
\frac{A_{A,nd}}{A_{A,d}}=1.4,
\end{equation}
by substituting \cref{A1} and \cref{A2} into the equation above we find that for Design A
\begin{equation} \label{DA}
D_{A,nd}=1.4\sqrt{2}D_{A,d}\approx 2 D_{A,d}.
\end{equation}
Substitution of \cref{DA} into \cref{V1} and \cref{V2} yields 
\begin{equation}
\frac{V_{A,nd}}{V_{A,d}}=\frac{2 L \pi D^2_{A,nd}}{2\sqrt{2}L \pi D^2_{A,d}}=2\sqrt{2}
\end{equation}
and
\begin{equation}\label{VT}
V_{A,T}=V_{A,nd}+V_{A,d}=2\pi L (D_{A,nd}^2+\sqrt{2} D_{A,d}^2)=2\pi L D_{A,nd}^2 \left(1+\frac{1}{2\sqrt{2}}\right),
\end{equation}
where $V_{A,T}$ indicates the total volume of the unit cell for Design A.

@@

\subsubsection{Rectangular cross-section}
From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ . Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
\begin{equation}
	V_{A,nd}=8LD_{A,nd}t
\end{equation}
\begin{equation}
	V_{A,d}=8\sqrt{2}LD_{A,d}t
\end{equation}
Let us enter this into our biological observation, which yields:
\begin{equation}
	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
\end{equation}
Therefore, we can simplify this relationship as
\begin{equation}
	D_{A,nd}=2D_{A,d}
\end{equation}

\subsection{Design B}
Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{DesignB}). However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. 
\subsubsection{Circular cross section}
It follows that the non-diagonal and diagonal volumes are given by
\begin{equation}
V_{B,nd}=V_{A,nd}=2\pi L D_{B,nd}^2
\end{equation}
and
\begin{equation}
V_{B,d}=2\sqrt{8} L \left(\pi \frac{{D}_{B,d}^2}{4}\right),
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2}, as well as the above volumes, we  obtain 
\begin{equation}
{{D}_{B,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{B,d}}{{D}_{B,nd}}=\frac{1}{\sqrt{2}}.
\end{equation}

@@

\subsubsection{Rectangular cross section}

For this design the volume of non-diagonal elements remain the same, namely,
\begin{equation}
		V_{B,nd}=8LD_{B,nd}t.
\end{equation}
However, the volume for the diagonal elements will is different as a result the change in total length, namely,
\begin{equation}
		V_{B,d}=4\sqrt{2}LD_{A,d}t.
\end{equation}
In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
\begin{equation}
		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
\end{equation}
which simplifies to
\begin{equation}
	D_{B,nd}=D_{B,d}
\end{equation}
and
\begin{equation}
	D_{B,nd}=D_{A,nd}
\end{equation}

\subsection{Design C}
Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{DesignC}).

\subsubsection{Circular cross section}
For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
\begin{equation}
V_{C,nd}=V_{A,nd}=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}
V_{C,d}=V_{A,d}=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2} we  obtain 
\begin{equation}
{{D}_{C,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{C,d}}{{D}_{C,nd}}=\frac{1}{2}.
\end{equation}

@@

\subsubsection{Rectangular cross section}
Because the total length of the diagonal and non-diagonal elements are the same as for Design A, we obtain the same diameters for the elements, namely,
\begin{equation}
D_{A,nd}=2D_{A,d}
\end{equation}
and
\begin{equation}
D_{C,nd}=D_{A,nd}
\end{equation}

\subsection{Design D} 
Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}. As such, for this design we allocate the total material volume to the non-diagonal elements.

\subsubsection{Circular cross section}
Since
\begin{equation}
V_{D,T}=V_{D,nd}=V_{A,nd}=2\pi L D_{D,nd}^2,
\end{equation}
using the constraint provided by \cref{con1} we obtain
\begin{equation}
{D_{D,nd}}={D_{A,nd}}\sqrt{1+\frac{\sqrt{2}}{4}}. 
\end{equation}

@@

\subsubsection{Rectangular cross-section}
For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements. Thus we can formulate the total volume of Design A as
\begin{equation}
	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
\end{equation}
Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
\begin{equation}
V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}
\end{equation}

@@

\section{Additional numerical results}
\KB{add here all additional numerical results}
In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value. This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.

@@

\section{Optimization Analysis}

\KB{what target functions are you using? Which parameters are you identifying via the optimization? How do you choose the initial values for all variables? Which value do you use for the initial standard deviation?}

For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}. The algorithm input variables are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction. Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements. For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.
